MinCall --- MinION end2end convolutional deep learning basecaller

### 2017-07-01 paper/main.tex
Added basecall speeds and removed input pipeline figure

We compared our model on R9 chemistry with Metrichor (HMM-based approach), Nanonet\footnote{\url{https://github.com/nanoporetech/nanonet/}} and DeepNano~\cite{deepnano} (RNN based approaches). For the detailed basecaller overview see the appendix~\ref{app:basecallers}.

@@

Dataset used were E.Coli K-12 strands from~\cite{loman1-100k} and Lamba basecalling\footnote{Acquired from doc. dr. sc. Petra Korać and dr.sc. Paula Dobrinić}. Both used datasets show in table \ref{tbl:datasets} have been previously passed through MinKNOW and had been basecalled by Metrichor. As 1D read analysis was the focus of this paper, only those reads were used.

@@

To eliminate the possibility of overfitting to the known reference, the model is trained and tested on reads from different organisms. Due to the limited amount of public available raw nanopore sequence data, E. Coli was \emph{divided} into two regions.
Reads were split into train and test portions, depending on which region of E. Coli they align.
If read aligns inside first 70\% of the E. Coli, it is placed into train set, and if it aligns to the second portion, it is placed into test set. Reads whose alignment overlaps train and test region are not used. Important to note that E. Coli genome, and genomes of the majority of other bacteria, is cyclical, so reads with alignments that wrap over edges are also discarded. Total train set consist of over 110 thousand reads.
Overview of the entire learning pipeline is shown in figure~\ref{fg:train_pipe}.

\begin{figure}[!ht]
    \begin{center}
        \includegraphics[width=0.7\textwidth]{./imgs/train_pipeline.png}
        \caption{Training pipeline overview}
        \label{fg:train_pipe}
    \end{center}
\end{figure}

@@

The fact that ground truth is not known makes evaluation difficult. Different methods for evaluation were used to get clearer information about each basecaller.

>>>>

We compared our model on R9 chemistry with Metrichor (HMM-based approach), Nanonet\footnote{\url{https://github.com/nanoporetech/nanonet/}} and DeepNano~\cite{deepnano} (RNN based approaches). For the detailed basecaller overview see the Appendix~\ref{app:basecallers}.

@@

Dataset used were E.Coli K-12 strands from~\cite{loman1-100k} and Lamba basecalling\footnote{Acquired from doc.~dr.~sc. Petra Korać and dr.sc. Paula Dobrinić}. Both used datasets show in table~\ref{tbl:datasets} have been previously passed through MinKNOW and had been basecalled by Metrichor. As 1D read analysis was the focus of this paper, only those reads were used.

@@

To eliminate the possibility of overfitting to the known reference, the model is trained and tested on reads from different organisms. Due to the limited amount of public available raw nanopore sequence data, E. Coli was \emph{divided} into two regions.
Reads were split into train and test portions, depending on which region of E. Coli they align.
If read aligns inside first 70\% of the E. Coli, it is placed into train set, and if it aligns to the second portion, it is placed into test set. Reads whose alignment overlaps train and test region are not used. Important to note that E. Coli genome, and genomes of the majority of other bacteria, is cyclical, so reads with alignments that wrap over edges are also discarded. Total train set consist of over 110 thousand reads.

% Overview of the entire learning pipeline is shown in figure~\ref{fg:train_pipe}.
% \begin{figure}[!ht]
%     \begin{center}
%         \includegraphics[width=0.7\textwidth]{./imgs/train_pipeline.png}
%         \caption{Training pipeline overview}
%         \label{fg:train_pipe}
%     \end{center}
% \end{figure}

@@

The fact that ground truth is not known makes evaluation difficult. Different methods for evaluation were used to get clearer information about each basecaller. In the following subsections are different evaluation metrics described and analysed. For completeness speed measurements are given in Table~\ref{tbl:speeds}

\begin{table}[]
	\caption{Base calling speeds measured in \textit{base pairs per second}. CPU is Intel Xeon E5-2640 v2 @ 2Ghz. GPU is NVIDIA Titan X Black}
	\label{tbl:speeds}
	\centering

	\begin{tabular}{lcc}
		\toprule
		{} &  \thead{ecoli (bp/s)} &  \thead{lambda (bp/s)} \\
		\midrule
		resdeep (CPU)  &               \textbf{1174.28} &              \textbf{ 1363.340 }\\
		Nanonet (CPU)  &                856.01 &                897.499 \\
		DeepNano (CPU) &                626.99 &                692.370 \\
		\midrule
		resdeep (GPU)  &              \textbf{ 6571.76 }&              \textbf{ 6140.300} \\
		Nanonet (GPU)  &               3828.39 &               3787.510 \\
		\bottomrule
	\end{tabular}

\end{table}
